Supplemental Materials

### 2021-06-22 supplement/supplement-main.Rmd
fixed prior commit

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.pos = 'H')
```

@@

The probability of a node Y can be defined from its parents X in many different ways. For binary nodes, the most flexible specification is a conditional probability table (CPT) with $2^n$ entries specifying the probability of Y given the entire joint probability distribution over X, where n is the number of parents X. Given the exponential complexity of the CPT, simpler functions are often used to specify the probability of the child node, such as the noisy-or and noisy-and-not functions. A noisy-or function assumes each of the parents are independent generative forces, each with a probability of generating the effect. A noisy-and-not function assumes each of the parents are independent preventive forces, specifying the probability of the child node as the probability that it is generated _and not_ prevented by any one of its possible preventers. We construct our cognitive model parameterization using a combination of these two functions, using a distribution sometimes referred to as a “DeMorgan gate” (Maaskant & Druzdel, 2008). In our model, the probability of the child node Y can be expressed as:

@@

Given measures of participants’ credences in each of the parent beliefs $x_i$ and the child belief $y$, we assume the following beta regression model:

$$ \mu = \exp\Bigg(\ln\Big(1 - \exp\big(\ln\big(1-w_0) + \sum_i(1-w_i)x_i\big)\big) \Big) + \ln\Big(\sum_i(1-w_i)(1-x_i)\Big)\Bigg) $$
$$y \sim beta\big(\mu k, (1-\mu)k\big)$$

Via maximum likelihood estimation, we estimate for each parent a discrete parameter $x_i$ determining whether the relationship is generative or preventive, and a continuous weight $w_i$ determining its strength. We also estimate $w_0$, the “leak probability” which is assumed to be generative, and _k_, a measure of dispersion in the responses _y_.

@@

One way to learn a Bayesian network structure from data is to search through a space of networks, “score” each and choose the best scoring network. Many scoring metrics are possible, but a directly meaningful one is p(d|G), the likelihood of the data given the graph structure. Asymptotically, this can be estimated using the Bayesian Information Criterion (BIC) (citation):

@@

Which we estimate using the maximum log-likelihood estimated for our statistical model, calculated as the sum of the maximum log-likelihood for each node given its parents, from a total of _k_ parameters and _n_ data points.

@@

Before constructing our model, we sorted the 14 measured beliefs into “tiers” based on how broad or abstract each belief was. For instance, we considered holistic balance and naturalism to be the most abstract beliefs among those we measured, and labeled these “worldviews”; we considered our outcome of interest, vaccine intentions, to be the most concrete measurement of a specific intended action. In between, we judged “medical skepticism”, “parental protectiveness”, and “parental expertise” to be more general than beliefs about diseases and vaccines, and so separated these into two tiers as well. We then applied the generative model principle to induce a partial-ordering of these beliefs, and a corresponding “blacklist” stipulating that certain edges must not appear in the final DAG. For instance, that there is no edge from vaccine intentions to naturalism. This constraint can be thought of as a strong prior over the possible structures of this intuitive theory (namely, P(G)=0 for any graph G not respecting the principle).

@@

To conduct an efficient search through the super-exponential space of DAGs, we tested several variants of two structure learning algorithms implemented in the “bnlearn” R package. The first is purely score-based “hill-climbing” algorithm (cite) and the second was a hybrid algorithm that combined the hill-climbing algorithm with a prior “restriction” phase that pruned down the space of DAGs. This pruning was accomplished with the “min-max parents and children” algorithm (cite), which can be tuned with a hyperparameter alpha, which we tested at several levels. To speed computations, we restricted the maximum number of parents in the graph to five for all nodes, as the complexity of the scoring function is exponential in the number of parents. In all resulting networks no node was found to have more than four parents, suggesting this did not impact the final results.

@@

library(tidyverse)
library(brms)
library(bnlearn)
library(HydeNet)

@@

```{r fit_all_models, include=FALSE}

# hc_args <- list(score="custom", fun=cog_model_score_func, maxp=5, max.iter=1000)
# 
# models <- list(
#   hc = partial(hc, score="custom", fun=cog_model_score_func, maxp=5, max.iter=1000),
#   hc_intentdv = partial(hc, blacklist=intent_dv_bl, score="custom", fun=cog_model_score_func, maxp=5, max.iter=1000),
#   hc_theory = partial(hc, blacklist=theory_bl, score="custom", fun=cog_model_score_func, maxp=5, max.iter=1000),
#   mmhc_05 = partial(mmhc, restrict.args=list(alpha=.05), maximize.args = hc_args),
#   mmhc_01 = partial(mmhc, restrict.args=list(alpha=.01), maximize.args = hc_args),
#   mmhc_intentdv_05 = partial(mmhc, blacklist = intent_dv_bl, restrict.args=list(alpha=.05), maximize.args = hc_args),
#   mmhc_intentdv_01 = partial(mmhc, blacklist = intent_dv_bl, restrict.args=list(alpha=.01), maximize.args = hc_args),
#   mmhc_theory_05 = partial(mmhc, blacklist = theory_bl, restrict.args=list(alpha=.05), maximize.args = hc_args),
#   mmhc_theory_01 = partial(mmhc, blacklist = theory_bl, restrict.args=list(alpha=.01), maximize.args = hc_args)
# )
# 
# models <- mclapply(
#     models, 
#     function(f){f(train)}, 
#     mc.cores = detectCores(logical=TRUE)
#   )
# 
# arc_counts <- map(models, ~nrow(arcs(.x)))
# arc_counts
```

@@

Without any clear guidance from cross validation and with a clear need to apply top-down constraints, we elected to advance our modeling with a model learned with the hill-climbing algorithm and our generative model constraints. This model contained more arcs than models estimated with hybrid algorithms, because these hybrid algorithms include a restriction phase based on correlations rather than our statistical model, we preferred the unrestricted approach.

@@

We used a bootstrapping procedure to estimate the confidence in each of the links in this model, conducting the structure learning procedure repeatedly for N bootstrapped resamples of the original training data. Figure X visualizes the frequency with which each arc appeared in the resulting model across these replicates, and figure x visualizes the model with the coefficients for each link encoded by color. [interpret the models]

@@

```{r}
## explore cpdag

@@

@@

```{r boot_strength_plot, echo=F, message=F, fig.cap="Bootstrapped confidence estimates for Bayesian network edges. Edge color indicates frequency with which edges appeared when conducting search across X bootstrap replicates. Only edges appearing in at least 20\\% of graphs are shown."}

hyde_str <- df_to_hyde_string(boot_res %>% filter(strength>.2, direction>0) %>% select(from,to))
h2 <- HydeNetwork(as.formula(hyde_str))
# plot(h2)

z2 <- custom_plot.HydeNetwork(h2)
boot_res %>%
  filter(direction>0, strength>.2) %>% 
  # filter(parameter!="phi",parameter!="w_leak") %>%
  # rename(from = parameter) %>%
  # mutate(from = gsub("w_","",from)) %>%
  mutate(strength = strength*direction) %>% 
  select(to, from, strength) %>%
  merge(z2$nodes_df %>% rename(from = label) %>% select(id, from), by="from") %>%
  mutate(from = id) %>%
  select(-id) %>%
  merge(z2$nodes_df %>% rename(to = label) %>% select(id, to), by="to") %>%
  mutate(to = id) %>%
  arrange(to,from) -> edge_df_mod

# z$nodes_df
z2$nodes_df <- z2$nodes_df %>% mutate(label = recode_nodes(label))
z2$nodes_df$fontname <- "Helvetica"
z2$nodes_df$shape <- "ellipse"
# z2$nodes_df$fillcolor <- ifelse(z2$nodes_df$id==13, "grey","white")
z2$nodes_df$fixedsize <- FALSE
z2$edges_df$penwidth <- abs(edge_df_mod$strength)*2
z2$edges_df$penwidth <- 2
# z$edges_df$label <- round(edge_df_mod$value,2)
# z$edges_df$color <- ifelse(edge_df_mod$value > 0, "blue","red")
z2$edges_df$color <- map_to_brewer(edge_df_mod$strength, bins=9, option="RdPu")
# z2$edges_df$color <- map_to_viridis(edge_df_mod$strength, option="magma")

render_graph(z2)
```

```{r}
boot_graph_legend <- (tibble(
  Probability = seq(min(edge_df_mod$strength), max(edge_df_mod$strength), length=11),
) %>% 
  ggplot(aes(y=1, x = Probability, fill = Probability)) +
  geom_tile() +
  scale_fill_distiller(palette="RdPu", direction=1)
) %>% 
  cowplot::get_legend()

cowplot::ggdraw(boot_graph_legend)
```

@@

```{r}
boot_df <- boot_res %>% filter(strength>.2, direction>0)
```

@@

export_graph(z, "../local/main-graph.pdf")
render_graph(z)

```
```{r}
coefs %>%
  filter(parameter!="phi",parameter!="w_leak") %>%
  rename(from = parameter) %>%
  mutate(from = gsub("w_","",from)) %>%
  select(to, from, value) %>%
  merge(z$nodes_df %>% rename(from = label) %>% select(id, from), by="from") %>%
  # mutate(from = id) %>%
  rename(from_id = id) %>% 
  merge(z$nodes_df %>% rename(to = label) %>% select(id, to), by="to") %>%
  rename(to_id = id) %>%
  arrange(to_id,from_id)
```

```{r}
## NOTE: I just can't figure out how to nicely display the graph and legend in code
## it's easier just to export them both and then combine with a graphics app

@@

@@

plt.valid
```
For each node, we also compared the Bayesian network model’s predictions to a maximal naive model using values on all other belief scales as predictors. This captures the degree to which each belief is predictable from the others as a whole. As shown in table X, the relatively simpler Bayesian network model does not sacrifice any predictive power compared to the suite of maximal naive models.

@@

kbl(model_fits_print, booktabs = T,
    digits = 3, 
    align = c('l','c','c','c'),
    # format = "latex",
    caption = "Correlations for observed and predicted out-of-sample values") %>% 
  kable_classic(full_width = F, html_font = "Cambria")
```

## Estimating the strength of evidence from changes in belief reports

To model the effects of participants observing evidence, we augmented the Bayesian network representing the cognitive model (shown in figure #) with an additional node representing the evidence. For instance, to model the effects of the “disease risk” intervention (Study 2), we added an evidence node as a child of “disease severity,” acting as virtual evidence for this node (following Pearl, 1988). To compute the CPT for this node as a function of disease severity beliefs, we estimated an _evidence ratio_ based on participants’ pretest and posttest credences in disease severity. To so do, we made use of the log-odds expression of Bayes rule,

$$ ln\big(O(h|d)\big) = ln \big(O(h)\big) + ln\Bigg(\frac{P(d|h=1)}{P(d|h=0)}\Bigg) $$

@@

The evidence ratio can be estimated from a model that is linear in the log-odds. This allowed for estimation of the evidence ratio from the with the following beta-regression model:

$$logit(\mu) = logit(pretest) + \ln(ER_0) + x\ln(ER_{1})$$
$$y \sim beta\big(\mu k, (1-\mu)k\big)$$

@@

$$ p(evidence = 1 | x ) = .50*ER_1$$ 
$$ p(evidence = 0 | x) = .50. $$

@@

## Qualitative Studies

@@

To do this, we drew on a variety of sources, including academic articles on anti-vaccine skepticism (e.g., CITE) and anti-vaccine websites, and a qualitative survey with self-identified vaccine skeptics (_n_=16) via Amazon Mechanical Turk (MTurk). [add details of this study here]

Ultimately, we generated a list of 13 underlying beliefs that might influence people’s decisions about whether to vaccinate their children (_vaccine intentions_, the last of our 14 beliefs). We translated these qualitative insights into 14 psychometrically-valid scales. In developing these scales, we stipulated that each scale should be brief, composed of 4-6 statements for participants to evaluate; include at least one reverse-coded item; and be highly reliable (Cronbach’s α ≥ .80). After extensive piloting and refinement, we created 14 scales that met these criteria, including one preexisting scale (the “holistic balance” subscale from McFadden et al., 2010).

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(fig.pos = 'H')
```

This project represents a line of theoretical and empirical work seeking to understand misconceptions and belief revision in the context of people’s intuitive theories. We focus our efforts on reducing vaccine skepticism by identifying people’s intuitive theories surrounding vaccination decisions. Through this case study we hope to shed light on why such misconceptions persist and how educators and interventionists might address misconceptions more effectively across domains. We argue that correcting misconceptions requires presenting evidence in a way that is sensitive to and persuasive within the broader conceptual context in which those misconceptions reside (see Gershman, 2018; Weisman & Markman, 2017). Understanding the revision of any one specific belief (e.g., a misconception like _vaccines are dangerous_) demands a holistic understanding of the wider set of related beliefs that make up people's intuitive theories.

First, we describe in further detail the modeling process for developing a Bayesian network model of the intuitive theory surrounding vaccination decisions.

Then, we describe our empirical methods in detail. This begins with the qualitative research we conducted with vaccine skeptics to develop psychometrically-valid quantitative scales measuring a network of 14 beliefs related to vaccine skepticism. Then, we report the detailed methods for each of the empirical studies reported in the paper. In Study 1, we used Bayesian network structure learning algorithms to develop a cognitive model of the relationships among these beliefs in a large sample of people. This model can predict how beliefs correlate across people and how they would be revised in the face of evidence. In Study 2, we replicated the success of an existing intervention emphasizing the dangers of childhood diseases (Horne et al., 2015) and used our model to provide a more detailed understanding of _how and why_ this intervention increases vaccine intentions. In Study 3, we drew inspiration from our model to create and validate the success of  a novel intervention aimed at debunking concerns about toxic additives. Finally, in Study 4 we tested whether people's beliefs changed coherently following relevant real-world events, by examining beliefs before and after serious outbreaks of measles in the United States in early 2019.

@@

In this graph, the edge $FIRE \rightarrow SMOKE$ captures that fire causes smoke.^[The example of cause and effect is especially intuitive, but these edges need not be causal. Instead, they are better thought of as representing “influence relations”, which may also include relationships like logical implication and set membership (Williamson, 2001).] This relationship is represented as a probabilistic function, where the probability of smoke depends on the probability of fire. The probability of a node Y can be defined from its parents X in many different ways. For binary nodes, the most flexible specification is a conditional probability table (CPT) with 2^n^ entries specifying the probability of Y given the entire joint probability distribution over X, where n is the number of parents X. Given the exponential complexity of the CPT, simpler functions are often used to specify the probability of the child node, such as the noisy-or and noisy-and-not functions. A noisy-or function assumes each of the parents are independent generative forces, each with a probability of generating the effect. A noisy-and-not function assumes each of the parents are independent preventive forces, specifying the probability of the child node as the probability that it is generated _and not_ prevented by any one of its possible preventers. We construct our cognitive model parameterization using a combination of these two functions, using a distribution sometimes referred to as a “DeMorgan gate” (Maaskant & Druzdel, 2008). In our model, the probability of the child node Y can be expressed as:

@@

This gives us a formalism for representing relations within an intuitive theory. Now we need methods for estimating the relationship that are actually represented within people’s intuitive theories. Given measures of participants’ credences in each of the parent beliefs $x_i$ and the child belief $y$, we assume the following beta regression model:

$$\mu = \exp\Bigg(\ln\Big(1 - \exp\big(\ln\big(1-w_0) + \sum_i(1-w_i)x_i\big)\big) \Big) + \ln\Big(\sum_i(1-w_i)(1-x_i)\Big)\Bigg)$$

$$y \sim beta\big(\mu k, (1-\mu)k\big)$$

Via maximum likelihood estimation, we estimate for each parent a discrete parameter $x_i$ determining whether the relationship is generative or preventive, and a continuous weight $w_i$ determining its strength. We also estimate $w_0$, the “leak probability” which is assumed to be generative, and $k$, a measure of dispersion in the responses $y$.

@@

Where the structure of an intuitive theory is known, Bayesian networks offer powerful tools for predicting and understanding how evidence will affect people's beliefs. However, there are many cases where the structure of the intuitive theory is unknown and is itself the very question under investigation.

Fortunately, the formalism of Bayesian Networks provides tools for navigating these challenges. Here, we used Bayesian network structure learning algorithms to construct a model of the intuitive theory of vaccine decisions (see Scutari & Denis, 2021). One way to learn a Bayesian network structure from data is to search through a space of networks, “score” each and choose the best scoring network. Many scoring metrics are possible, but a directly meaningful one is p(d|G), the likelihood of the data given the graph structure. Asymptotically, this can be estimated using the Bayesian Information Criterion (BIC) (citation):

@@

Which we estimate using the maximum log-likelihood estimated for our statistical model, calculated as the sum of the maximum log-likelihood for each node given its parents, from a total of $k$ parameters and $n$ data points.

@@

Before constructing our model, we sorted the 14 measured beliefs into “tiers” based on how broad or abstract each belief was. For instance, we considered holistic balance and naturalism to be the most abstract beliefs among those we measured, and labeled these “worldviews”; we considered our outcome of interest, vaccine intentions, to be the most concrete measurement of a specific intended action. In between, we judged “medical skepticism”, “parental protectiveness”, and “parental expertise” to be more general than beliefs about diseases and vaccines, and so separated these into two tiers as well. We then applied the generative model principle to induce a partial-ordering of these beliefs, and a corresponding “blacklist” stipulating that certain edges must not appear in the final DAG. For instance, that there is no edge from vaccine intentions to naturalism. This constraint can be thought of as a strong prior over the possible structures of this intuitive theory (namely, $P(G)=0$ for any graph $G$ not respecting the principle).

@@

To conduct an efficient search through the super-exponential space of DAGs, we tested several variants of two structure learning algorithms implemented in the “bnlearn” R package. The first is a purely score-based “hill-climbing” algorithm (cite) and the second was a hybrid algorithm that combined the hill-climbing algorithm with a prior “restriction” phase that pruned down the space of DAGs. This pruning was accomplished with the “min-max parents and children” algorithm (cite), which can be tuned with a hyperparameter alpha, which we tested at several levels. To speed computations, we restricted the maximum number of parents in the graph to five for all nodes, as the complexity of the scoring function is exponential in the number of parents. In all resulting networks no node was found to have more than four parents, suggesting this did not impact the final results.

@@

library(tidyverse)
library(brms)
library(bnlearn)
library(HydeNet)
library(parallel)

@@

@@

Without any clear guidance from cross validation and with a clear need to apply top-down constraints, we elected to advance our modeling with a model learned with the hill-climbing algorithm and our generative model constraints. This model contained more arcs than models estimated with hybrid algorithms, but because these hybrid algorithms include a restriction phase based on correlations rather than our statistical model, we preferred the unrestricted approach.

@@

We used a bootstrapping procedure to estimate the confidence in each of the links in this model, conducting the structure learning procedure repeatedly for 200 bootstrapped resamples of the original training data. Figure X visualizes the frequency with which each arc appeared in the resulting model across these replicates, and figure x visualizes the model with the coefficients for each link encoded by color.

@@

Though we found the uncovered network structure to be quite revealing and informative, there are also some important limitations that should be acknowledged. In particular, the edges from the vaccine danger and vaccine effectiveness nodes to the vaccine toxicity node strike us as counterintuitive. Intuitively, we might imagine that vaccines having toxins could cause them to be dangerous and ineffective. Yet the edges in the model run from in the opposite direction, and so do not support a clear causal interpretation.

We can examine the robustness of the model’s features in two ways. First, we can ask about the “skeleton” of the graphical model: whether the relationships between pairs of nodes are robust, irrespective of their directions. To test this, we applied our chosen structure learning algorithm against 1000 bootstrapped replicates of our data (following Friedman et al., 2013). As shown in figure X, we have a high degree of confidence in the “skeleton” of the graph: nodes which are connected in the final graph are reliably connected across bootstrapping runs.

@@

```{r, echo=F}
## examine bootstrapping results

## which arcs are most likely?
boot_res %>% 
  filter(strength > .80, direction > .5)

## what is strength of each arc in final graph?
bnlearn_to_df(main_model) %>% 
  left_join(boot_res, by=c("from","to"))  %>% 
  filter(strength > .80)
```

@@

```{r boot_strength_plot, echo=F, message=F, fig.cap="Bootstrapped confidence estimates for Bayesian network edges. Edge color indicates frequency with which edges appeared when conducting search across X bootstrap replicates. Only edges appearing in at least 20\\% of graphs are shown."}
# 
# hyde_str <- df_to_hyde_string(boot_res %>% filter(strength>.2, direction>0) %>% select(from,to))
# h2 <- HydeNetwork(as.formula(hyde_str))
# # plot(h2)
# 
# z2 <- custom_plot.HydeNetwork(h2)
# boot_res %>%
#   filter(direction>0, strength>.2) %>% 
#   # filter(parameter!="phi",parameter!="w_leak") %>%
#   # rename(from = parameter) %>%
#   # mutate(from = gsub("w_","",from)) %>%
#   mutate(strength = strength*direction) %>% 
#   select(to, from, strength) %>%
#   merge(z2$nodes_df %>% rename(from = label) %>% select(id, from), by="from") %>%
#   mutate(from = id) %>%
#   select(-id) %>%
#   merge(z2$nodes_df %>% rename(to = label) %>% select(id, to), by="to") %>%
#   mutate(to = id) %>%
#   arrange(to,from) -> edge_df_mod
# 
# 
# # z$nodes_df
# z2$nodes_df <- z2$nodes_df %>% mutate(label = recode_nodes(label))
# z2$nodes_df$fontname <- "Helvetica"
# z2$nodes_df$shape <- "ellipse"
# # z2$nodes_df$fillcolor <- ifelse(z2$nodes_df$id==13, "grey","white")
# z2$nodes_df$fixedsize <- FALSE
# z2$edges_df$penwidth <- abs(edge_df_mod$strength)*2
# z2$edges_df$penwidth <- 2
# # z$edges_df$label <- round(edge_df_mod$value,2)
# # z$edges_df$color <- ifelse(edge_df_mod$value > 0, "blue","red")
# z2$edges_df$color <- map_to_brewer(edge_df_mod$strength, bins=9, option="RdPu")
# # z2$edges_df$color <- map_to_viridis(edge_df_mod$strength, option="magma")
# 
# render_graph(z2)
```

```{r, echo=F, warning=F, message=F}
# boot_graph_legend <- (tibble(
#   Probability = seq(min(edge_df_mod$strength), max(edge_df_mod$strength), length=11),
# ) %>% 
#   ggplot(aes(y=1, x = Probability, fill = Probability)) +
#   geom_tile() +
#   scale_fill_distiller(palette="RdPu", direction=1)
# ) %>% 
#   cowplot::get_legend()
# 
# cowplot::ggdraw(boot_graph_legend)
```

Next, we can ask about the directions of the relationships within the model by examining the “completed partially directed acyclic graph” (CPDAG) derived from the structure learning algorithm. This is a graph with both directed and undirected edges which represents the “Markov equivalence class” of DAGs: the set of DAGs representing equivalent patterns of conditional (in)dependence among variables. Alternatively, this can be conceived of as “score equivalence,” the set of graphs which would produce equivalent scores for the observed data (Chickering, 2013). The CPDAG derived from our fit model contains 14 directed edges and 15 undirected arcs—indicating that the data do not constrain the directions of 15 of the edges in the final model. Of these undirected arcs, 10 directions are determined by our generative model constraints that we applied before fitting the model. Five arcs were determined neither by these constraints nor the data: these were $\text{holistic balance} \leftrightarrow \text{naturalism}$; $\text{medical skepticism} \leftrightarrow  \text{parental expertise}$, $\text{vaccine toxicity} \leftrightarrow \text{vaccine efficacy}$, $\text{vaccine toxicity} \leftrightarrow \text{vaccine danger}$ and $\text{vaccine danger} \leftrightarrow \text{vaccine efficacy}$. Thus, we feel that no strong conclusions should be drawn surrounding the counterintuitive relationships regarding vaccine toxicity, danger, and effectiveness.

```{r, include=F}
## explore cpdag

@@

@@

## export graph code as txt
graph_text <- DiagrammeR::generate_dot(z)
write_file(graph_text, "graph-dot-code.txt")

# export graph image and render
export_graph(z, "../local/main-graph.pdf")
render_graph(z)
```

```{r, echo=F, warning=F, message=F}
## NOTE: I just can't figure out how to nicely display the graph and legend in code
## it's easier just to export them both and then combine with a graphics app

@@

Altogether, these findings might be seen as demonstrating the need for a heavier hand in establishing top-down constraints on the model fitting. Certainly, they call for caution in interpreting the relationships represented in this model. Due to their tight connections with structural causal networks, it is tempting to interpret the edges in Bayesian networks as “causal”. However, these relationships are best thought of simply as probabilistic dependencies or “influence” relations (cite). Indeed, there are other cases where it is unclear if an edge should be directed at all: e.g., it is not obvious to us what would be the “correct” direction for the edge between naturalism and holistic balance. We might even speculate there is an additional unmeasured variable that connects these beliefs and that would best account for their association. Ultimately, this model should be understood as offering a plausible, but not definitive, account for a cognitive model connecting these different beliefs.

What does the resulting model reveal about the intuitive theory undergirding vaccination decisions?

The three nodes with direct connections to _vaccine intentions_ are beliefs about _vaccine effectiveness_, _vaccine danger_, and the severity of childhood diseases (_disease severity_)--concrete beliefs that are, at face value, especially closely related to vaccination decisions. The directionality of these edges is informed by the constraints we applied via the generative model principle, but the identification of these three factors as the most proximal to vaccine intentions is a product of the structure learning approach. This gives us some confidence that the model is capturing important relationships.

Other parts of the model shed new light on the role of lay theories in vaccine decisions. For example, _naturalism_—the general view that natural things are better than artificial things—appears to be strongly related to _medical skepticism_ and _parental expertise_; all three of these abstract beliefs are related to concrete beliefs that, in turn, feed into participants’ vaccination intentions. This hierarchical structure is at least partly a product of our constraints that sorted these beliefs into “tiers” of abstractness. However, the [fact that a model with these constraints still functions well for prediction demonstrates that this hierarchical structure is plausible].

Finally, this model structure also helps explain some initially surprising findings. For example, before collecting these data we speculated that dispelling the misconceptions about the capacity of the infant immune system could promote positive attitudes toward vaccination. We were disappointed to observe the weak first-order correlation between this belief and vaccine intentions in our behavioral data (r = -.097 in our training split). The model sheds light on this surprising (lack of) relationship: Although the belief that the infant immune system is limited in capacity is positively related to the belief that vaccines strain the immune system—discouraging vaccination, as we had assumed—it also seems to promote the belief that childhood diseases have severe consequences for young children (_disease severity_), which might, in turn, encourage vaccination. These countervailing forces—which, on average, seem to have cancelled each other out in this sample—dissuaded us from our plans to develop an intervention aimed at dispelling misconceptions about limited capacity in order to encourage childhood vaccination.

@@

plt.valid
```
For each node, we also compared the Bayesian network model’s predictions to a naive “maximal” model using values on all other belief scales as predictors. Our intention was to capture the degree to which each belief is predictable from the others as a whole. As shown in table X, the relatively simpler Bayesian network model does not sacrifice any predictive power compared to the suite of maximal naive models.

@@

kbl(model_fits_print, booktabs = T,
    digits = 3, 
    align = c('l','c','c','c'),
    # format = "latex",
    caption = "Correlations for observed and predicted out-of-sample values") %>% 
  kable_classic(full_width = F, html_font = "Cambria")
```
## Estimating the strength of evidence from changes in belief reports

Returning to the DAG shown in Figure 1 above, consider the subgraph $FIRE \rightarrow SMOKE$, capturing the fact that fire causes smoke. This relationship is represented as a probabilistic function, where the probability of SMOKE depends on the probability of FIRE. If we observe the presence of SMOKE, we can draw inferences about FIRE.

However, sometimes reasoners have only indirect evidence at their disposal. For instance, a person might not directly observe smoke, but instead hear from a witness who claimed to have seen it. Indirect or partial observations can be captured by augmenting the network, creating the graph $FIRE \rightarrow SMOKE \rightarrow WITNESS$. Here, a new node WITNESS represents another person’s testimony that they saw smoke, providing imperfect evidence for SMOKE, encoded by another probabilistic function. Observing WITNESS informs beliefs about SMOKE, in turn informing beliefs about FIRE. This approach to representing uncertain evidence is known as “virtual evidence” (Pearl, 1988).

To model the effects of participants observing evidence, we augmented the Bayesian network representing the cognitive model (shown in figure #) with an additional node representing the evidence. For instance, to model the effects of the “disease risk” intervention (Study 2), we added an evidence node as a child of “disease severity,” acting as virtual evidence for this node. To compute the CPT for this node as a function of disease severity beliefs, we estimated an _evidence ratio_ based on participants’ pretest and posttest credences in disease severity. To so do, we made use of the log-odds expression of Bayes rule,

$$ log\big(O(h|d)\big) = log\big(O(h)\big) + ln\Big(\frac{P(d|h=1)}{P(d|h=0)}\Big) $$

@@

The evidence ratio can be estimated from a model that is linear in the log-odds. This allows for estimation of the evidence ratio from the with the following beta-regression model:

$$logit(\mu) = logit(pretest) + \ln(ER_0) + x\ln(ER_{1})$$

$$y \sim beta\big(\mu k, (1-\mu)k\big)$$

@@

$$ p(evidence = 1 | x ) = .50*ER_1$$

$$ p(evidence = 0 | x) = .50. $$

@@

## Qualitative Study

@@

To do this, we drew on a variety of sources, including a qualitative survey with participants in a previous study who had self-identified as vaccine skeptics. Out of 23 people invited to participate, n=16 participants completed the survey via Amazon Mechanical Turk (MTurk). All participants answered “yes” to the question “Are you concerned that childhood vaccinations might cause serious side-effects?” When asked, “Do you have concerns about all vaccines or only some specific vaccines?” nearly all participants (14 of 16) expressed being concerned about specific vaccines; 3 of these participants, as well as the remaining 2 participants, expressed concern about all vaccines.

The remainder of the survey was devoted to open-ended questions, which we will summarize here as they were a major source of inspiration for the development of the 14 “belief scales” employed in Studies 1-4.

When asked about the potential side effects of vaccines, most participants (10 of 16) mentioned autism as a potential side effect; some mentioned other serious medical issues (e.g., stunting, cancer; 8 out 16); temporary, less serious physical symptoms (e.g., fever, headache; 7 of 16); other cognitive impairments and developmental disorders (e.g., learning disabilities, loss of speech; 6 of 16), and risk of death (6 of 16). Interestingly, 2 participants (of 16) specifically discussed the possibility that vaccines might have negative impacts on a child’s immune system; for example, one wrote that a vaccine “penetrates and actually weakens the immune system overall (despite increasing immunity to specific sicknesses).” In explaining _how_ vaccines might cause these side effects, many participants (9 of 15; 1 participant did not provide explanations) expressed concern that children’s bodies and immune systems may be too weak or underdeveloped to tolerate vaccines (e.g., “The vaccines are given to the children when they are babies and their brains have not begun development”; “Too many are given all at once at too young an age”; “They overwhelm the child's nervous system”). Several participants (5 of 15) mentioned the presence of specific toxic additives in vaccines (e.g., “high levels of heavy metal, especially mercury”), or invoked the idea that vaccines are generally unnatural (e.g., “Vaccines are unnatural additions to the body. The human body should naturally fight off the disease and we should not be controlling how nature plays its role on mankind”; “These vaccines are simply chemicals that should not play a role in a young naturally developing child”). Several participants (4 of 15) expressed concern that individual differences across children might render some children particularly vulnerable to these side effects (e.g., “The vaccine itself could interact with the specific child in a way that is unique and unforeseen”; “Sorry, I don't know the EXACT cause of it. I just think it has to do with the child's body being able to take the vaccine or not”). Individual participants also related personal stories of neighbors and friends, mentioned their mistrust of the government, or mentioned that the side effects of vaccines are worse than the symptoms of the diseases they are intended to prevent.

When asked whether vaccines were effective in preventing the diseases they were designed to prevent, participants were evenly split (8 said “yes” and 8 said “no”) -- but in both groups, participants’ open-ended explanations touched on a common set of caveats and concerns. For example, some participants (7 of 15; 1 participant did not provide explanations) indicated that some vaccines are effective, at least for some individuals, but other vaccines are not (e.g., “Some work for some people, some don't”). Several participants (5 of 15) indicated that, regardless of efficacy, vaccines are no longer necessary because the diseases they prevent are no longer prevalent (e.g., “In certain circumstances they are, but in many cases the disease isn't really much of a risk so it doesn't make sense to fill the body with chemicals”; “I think that the potential of getting a disease is overblown and people get scared so they needlessly get vaccines”). Two (of 15) of participants mentioned that vaccines would be effective if they were given to children on a modified schedule (e.g., “Yes, if you give them ONE AT A TIME, spaced out over time and starting at older ages”); and two (of 15) related their personal mistrust of pharmaceutical companies (e.g., “THEY SEEMED TO BE A BIG MONEY MAKING SCAM BY THE BIG PHARMS”). (Individual participants related additional personal theories and concerns.)

When asked whether it would be safer to vaccine a 2-month-old child for 5 separate diseases in one doctor’s visit (as the CDC recommends) or to spread these vaccinations out at a rate of 1 every 2 months, nearly all participants (15 of 16) said it would be safer to spread the vaccinations out. Many of these participants (8 of 15) indicated that this would be safer because it would prevent the  immature system from being overwhelmed (e.g., “It gives a [child’s] body time to adapt and doesn't overwhelm their tiny immune systems”); and several (4 of 15) suggested that this would be preferable because it would allow the parent more control over the process (e.g., “Ideally I would choose neither for my child, but if forced to choose between the two I would pick the latter.  This would allow me to stop the vaccination process if things are not progressing well”; “I think a parent should have a choice which ones to have given to their child”). At least two participants expressed being deeply appalled by this choice, e.g., “He is only 8 weeks old. Who in their right mind decided it was safe to put 5 disease substances in a baby? Dead or alive, still horrific”; Honestly I do not think either is safe I think in both instances the child's body and health is actually being raped.”

Finally, when asked why most healthcare providers recommend vaccinating young children, most participants (7 of 15; 1 participant did not respond) stated explicitly that healthcare providers have financial incentives to recommend vaccination -- but many (7 of 15, including several who mentioned financial incentives) also acknowledged that healthcare providers likely believe (in the participants’ view, mistakenly) that vaccinations are harmless and effective. Some participants (6 of 15) stated that healthcare providers are following protocol

Drawing on these responses as well as anti-vaccine websites and academic articles on anti-vaccine skepticism (e.g., CITE), we generated a list of 13 underlying beliefs that might influence people’s decisions about whether to vaccinate their children (_vaccine intentions_, the last of our 14 beliefs). These included a variety of claims about vaccines, including beliefs about (1) the overall danger of vaccines (_vaccine danger_); (2) _toxic additives in vaccines_; and (3) _vaccine effectiveness_, how effective vaccines are in preventing disease; as well as a variety of specific claims about childhood diseases like measles, mumps, and rubella, including beliefs about (4) _disease rarity_ and (5) _disease severity_. Inspired by the many participants who invoked concerns about the effect of vaccines on infants’ immature immune systems, we also developed scales to assess beliefs that (6) the infant immune system is weak (_IIS: weakness_); (7) the infant immune system is limited in its capacity and can be easily overwhelmed (_IIS: limited capacity_); and (8) vaccines strain the infant immune system (_IIS: vaccines strain_). Beyond this, we developed two scales to assess general theories about parenting, including (9) general _parental protectiveness_; (10) _parental expertise_, namely the belief that parents usually know more about their children’s health than medical experts; and one scale to assess (12) _medical skepticism_, including concerns about pharmaceutical companies and corruption in the medical community. Finally, we decided to assess two broad worldviews: (12) _naturalism_, a general preference for natural over artificial things; and (13) _holistic balance_, one important aspect of attitudes toward alternative medicine (e.g., “The body is essentially self-healing and the task of a health care provider is to assist the healing process”; McFadden, Hernandez, & Ito, 2010).

_Holistic balance_ was the only belief for which we could identify a pre-existing psychometrically valid scale (McFadden et al., 2010). For each of the other 13 beliefs, we translated our qualitative insights into new psychometrically-valid scales through an iterative series of pilot studies not described here. We stipulated that each scale should be brief, composed of 4-6 statements for participants to evaluate; include at least one reverse-coded item; and be highly reliable (Cronbach’s α ≥ .80).
